\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)

### 2015-09-20 overview.tex
updates in section 3

\textbf{Safety specifications}.
In addition to the usual assertions that can be used to write safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special machine that can receive, but not send, events. A safety monitor maintains local state that is modified in response to events received from ordinary (non-monitor) machines. This local state is used to maintain a history of the computation that is relevant to the property being specified. An erroneous global behavior is flagged via an assertion on the private state of the safety monitor. 
Thus, a monitor cleanly separates instrumentation state required for specification (inside the monitor) from program state (outside the monitor).

To declare a monitor, the programmer inherits from the \psharp \texttt{Monitor} class. Monitors in \psharp are singleton instances, which means that an ordinary machine does not need a reference to a monitor to send it an event. An ordinary machine can send an event to a monitor by calling \texttt{Monitor<M>(e, p)}, where the parameter \texttt{e} is the event being send together with an optional payload \texttt{p}.

\textbf{Liveness specifications}.
Liveness property specifications are more difficult to encode since they are intended to capture program progress
over infinite executions.
Violation of a liveness property can only be demonstrated by an infinite execution.
Usually, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977,lamport1994temporal}.
We take a different approach and allow the programmer to write a \emph{liveness monitor}.
Similar to a safety monitor, a liveness monitor receives events but does not send them.
Unlike a safety monitor, a liveness monitor~\cite{desai2015building}
contains three types of states---\emph{ordinary}, \emph{hot}, and \emph{cold}.
A hot state denotes a point in the execution where progress is required but has not happened yet,
e.g., a node has failed but a new one has not come up yet.
A liveness monitor enters a hot state upon receiving notification of an event that requires the system
to make progress and leaves the hot state upon receiving another event notifying it of progress.
An infinite execution is erroneous if the liveness monitor is in a hot state infinitely often but in a cold state only finitely often.
Our liveness monitors can encode arbitrary temporal logic properties.

A thermometer analogy is useful for understanding a liveness monitor.
A hot state increases the temperature by a small value.
A cold state resets the temperature to zero.
A liveness error happens if the temperature becomes infinite.
%The distinction between ordinary and cold states is useful for modeling progress for each instance of an infinite stream of events,
%such as a periodic request or a periodic timer expiration.

A liveness violation is witnessed by an infinite execution in which all concurrently executing \psharp machines are scheduled fairly.
Unfortunately, it is clearly not possible to generate an infinite execution by executing a program for a finite amount of time.
Therefore, our implementation of liveness checking approximates an infinite execution using several heuristics.
One heuristic considers an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}.
Another heuristic maintains a cache of (pieces of) the state of the \psharp program obtained at each step in the execution
and reports an ``infinite'' execution when the latest step results in a cache hit, thus approximating a cycle in the state
graph of the program.

>>>>

\textbf{Safety specifications}.
In addition to the usual assertions for stating safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special machine that can receive, but not send, events.

A safety monitor maintains local state that is modified in response to events received from ordinary (non-monitor) machines. This local state is used to maintain a history of the computation that is relevant to the property being specified. An erroneous global behavior is flagged via an assertion on the private state of the safety monitor. Thus, a \psharp monitor cleanly separates instrumentation state required for specification (inside the monitor) from program state (outside the monitor).

\textbf{Liveness specifications}.
Liveness property specifications are more difficult to encode since they are intended to capture program progress over infinite executions. Violation of a liveness property can only be demonstrated by an infinite execution. Usually, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977,lamport1994temporal}. We take a different approach and allow the programmer to write a \emph{liveness monitor}~\cite{desai2015building}. Similar to a safety monitor, a liveness monitor can receive, but not send, events. Unlike a safety monitor, a liveness monitor contains two types of states: \emph{hot} and \emph{cold}.

A hot state denotes a point in the execution where progress is required but has not happened yet, for example a node has failed but a new one has not come up yet. A liveness monitor enters a hot state upon receiving notification of an event that requires the system to make progress. The liveness monitor leaves the hot state and enters a cold state upon receiving another event notifying it of progress. An infinite execution is erroneous if the liveness monitor is in a hot state infinitely often but in a cold state only finitely often. Our liveness monitors can encode arbitrary temporal logic properties. A thermometer analogy is useful for understanding liveness monitors: a hot state increases the temperature by a small value; a cold state resets the temperature to zero; a liveness error happens if the temperature becomes infinite.
%The distinction between ordinary and cold states is useful for modeling progress for each instance of an infinite stream of events, such as a periodic request or a periodic timer expiration.

A liveness violation is witnessed by an \emph{infinite} execution in which all concurrently executing \psharp machines are \emph{fairly} scheduled. Unfortunately, it is clearly not possible to generate an infinite execution by executing a program for a finite amount of time. Therefore, our implementation of liveness checking in \psharp approximates an infinite execution using several heuristics. One such heuristic considers an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}. Another heuristic maintains a cache of (pieces of) the state of the \psharp program obtained at each step in the execution, and reports an ``infinite'' execution when the latest step results in a cache hit, thus approximating a cycle in the state graph of the program.

\textbf{Using monitors}.
To declare a safety or a liveness monitor, the programmer must inherit from the \psharp \texttt{Monitor} class. \psharp monitors are singleton instances and, thus, an ordinary machine does not need a reference to a monitor to send it an event. Instead, an ordinary machine can invoke a monitor by calling \texttt{Monitor<M>(e, p)}, where the parameter \texttt{e} is the event being send, and the parameter \texttt{p} is an optional payload.
